---
title: "Bogota car accidents 2014"
output:
  github_document: default
  html_notebook: default
---

# Bogota car accidents 2014

```{r setup}
library(tidyverse)
library(lubridate)
```


## Data import

Data downloaded from https://docs.google.com/spreadsheets/d/1lgR-UpR9v9lmtoeK1ZhlVfs0FBj1xXz_eImhcrSI4Q0/

```{r}
accidents_raw <- read_csv("data/Car_Accidents_2014_Bogota.csv")
accidents_raw
```

* `the_geom` <chr> : geom
* `cartodb_id` <int>        ???
* `codigo_accidente` <int>  ??? : Accident code from police ?
* `fecha_ocurrencia` <date>       2014-10-29 : Date
* `hora_occurencia` <chr>   ???   14:10:00 : Time
* `gravedad` <chr> : Gravity (severity)
* `objectid` <int>          ???
* `localidad` <chr> : Location
* `municipio` <chr> : Municipality
* `x` <dbl> : Longitude
* `y` <dbl> : Latitude
* `timedate` <S3: POSIXct>        2014-10-29 14:10:00
* `clase` <chr> : Class [choque = shock; atropello = ; otro = ; caida de ocupante = ]




```{r}
# The data is for one year (2014). If we want to filter by month (etc), need a `month` variable first. Could use **lubridate**, or `separate()` ...The former avoids 10/11/12 coming before 2/3/4/etc, but the latter should also be ok in plots if using integers

?separate

accidents <- accidents_raw %>%
  separate(fecha_ocurrencia, c("year", "month", "day"), "-", convert = TRUE) %>%
                                              # `convert` gets y&m&d as integers
    filter(year == 2014)                      # 17 rows from 2013 filtered out

accidents <- accidents %>%
  separate(hora_ocurrencia, c("hour", "minute", "second"), ":", convert = TRUE) %>%
    filter(hour != "1899-12-30")              # 9 rows filtered out

accidents                                     # (9973 rows x 17 cols)
```


## Basic EDA

```{r}
# Just to get a sense of what's going on

accidents %>% count(the_geom)            # 6921 rows (so not unique)
accidents %>% count(the_geom) %>% count(n)  # 21 rows (1 to 22, without 18)
accidents %>% count(cartodb_id)          # 9973 rows (confirmed as unique)
accidents %>% count(codigo_accidente)    # 9973 rows (unique). Seem to be 5e+05 or 4e+06
accidents %>% count(month)               # 12 rows. `n` is 485-1584. Higher in Oct/Nov/Dec
accidents %>% count(hour)                # 24 rows. `n` is 51-722 (2am, 2pm)
accidents %>% count(gravedad)            # 3 rows. "CON HERIDOS", "CON MUERTOS", "SOLO DANOS"
accidents %>% count(objectid)            # 9973 rows (confirmed as unique)
accidents %>% count(localidad)           # 19 rows. `n` is 52-1176
accidents %>% count(municipio)           # 1 row ("BOGOTA"). `n` is 9973
accidents %>% count(clase)               # 7 rows. `n` is 1 ("INCENDIO") to 8494 ("CHOQUE")
```

Initial data visialisation, to see if there might be a location/ time pattern

```{r}
accidents <- accidents %>%
  mutate(date = make_date(year, month, day), week_no = week(date), week_day = wday(date, label = TRUE), hour = as.numeric(hour), n = n())

accidents

ggplot(accidents, aes(date)) +
  geom_histogram() +
  facet_wrap(~ localidad)

ggplot(accidents, aes(x, y, colour = localidad)) +
  geom_point()
```


Maybe look at a time-based model (especially if there is data for other years too, in which case start with localidad=ENGATIVA). And/or a visualisation (like ACotgreave? if I can get week#). Or time of day (model/viz). Or connected to streetlights (if there's a joining field)

```{r}
accidents2 <- accidents %>%
  group_by(week_no, week_day) %>%
  summarise(n = n())

accidents2

ggplot(accidents2, aes(week_day, week_no)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5))

ggplot(accidents2, aes(week_no, week_day)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5))
```


```{r}
accidents3 <- accidents %>%
  group_by(week_day, hour) %>%
  summarise(n = n())

accidents3

ggplot(accidents3, aes(week_day, hour)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5))

# Create a factor, to order the hours correctly (0-23) on the plot axis
# library(forcats)

# hour_levels <- c("0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23")

# y1 <- factor(hour, levels = hour_levels)
# Couldn't fix an error here. Instead converted `hour` in `accidents` from <chr> to numeric
# y1

# ggplot(accidents3, aes(week_day, y1)) +
#   geom_point(aes(colour = n, size = n, alpha = 0.5))

#--

# ggplot(accidents3, aes(hour, n)) +
#   geom_density(aes(colour = week_day)) +
#   facet_wrap(~week_day)
# Error in eval(substitute(list(...)), `_data`, parent.frame()) : object 'y' not found
```


```{r}
accidents4 <- accidents %>%
  group_by(month, week_day, hour, localidad) %>%  # error: grouping gives rings on viz not total
  summarise(n = n())

accidents4

ggplot(accidents4, aes(week_day, hour)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5)) +
    facet_wrap(~ localidad)

ggplot(accidents4, aes(week_day, hour)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5)) +
    facet_wrap(~ month)

ggplot(accidents4, aes(week_day, hour)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5)) +
    facet_grid(localidad ~ month)

ggplot(accidents4, aes(week_day, hour)) +
  geom_point(aes(colour = n, size = n, alpha = 0.5)) +
    facet_grid(localidad ~ month)
```


Try to create model of variation by hour. Start with localidad=ENGATIVA

```{r}
accidents5 <- accidents %>%
  group_by(hour) %>%
  summarise(n = n())

accidents5

ggplot(accidents5, aes(hour, n)) +
  geom_line()

ggplot(accidents3, aes(hour, n)) +
  geom_line(aes(colour = week_day))

ggplot(accidents3, aes(hour, n)) +
  geom_line(aes(colour = week_day)) +
  facet_wrap(~week_day)

# ggplot(accidents4, aes(hour, n)) +
#  geom_line(aes(colour = week_day)) +
#  facet_grid(. ~ week_day)
# Error in combine_vars(data, params$plot_env, cols, drop = params$drop) : At least one layer must contain all variables used for facetting
```



